### 2020-09-15 paper/coloringBergeGraphs.tex
Style fixes
Small fixes

A natural problem for perfect graphs is a problem of coloring them. In 1988 Gr\"otschel et al. published an ellipsoid-method-based polynomial algorithm for coloring perfect graphs \cite{Grtschel1993}. We consider it in \cref{sec:coloringEllipsoid}. However due to its use of the ellipsoid method this algorithm has been usually considered unpractical \cite{coloringSquareFree,Chudnovsky2003, coloringArtemis}.

There has been much progress on the quest of finding a more classical algorithm coloring perfect graphs, without the use of ellipsoid method (see \cref{sec:classicalColoring}), however there is still no known polynomial combinatorial algorithm to do this. \todo{better wording of this paragraph}

@@

\Cref{sec:InformationTheory} and \cref{sec:computingTheta} are based on lecture notes by Lovász \cite{Lovasz95}.

@@

Unfortunately, it is not known whether $\Theta(G)$ can be computed for all graphs in finite time. If we could calculate $\alpha(G^k)$ for a first few values of $k$ (we will show how to do it in \cref{alg:maxStableSet}) we could have a lower bound on $\Theta(G)$. Let us now turn into search for some usable upper bound.

@@

\begin{theorem}[Lovász "sandwich theorem" \cite{Knuth1994}]
  \label{thm:sandwich}
  For any graph $G$:
  $$ \omega(G) \leq \vartheta(\overline{G}) \leq \chi(G) $$
\end{theorem}

@@

First, let us recall some definitions, with \cite{gilbertstrang2020} as a reference for linear algebra.

@@

We have equivalent definitions of semidefinite matrices.
\begin{theorem}
  For a real symmetric $n \times n$ matrix $A$, the following are equivalent:
  \begin{enumerate}[(i)]
    \item $A$ is positive semidefinite,
    \item \label{en:ei2} for every $x \in \mathbb{R}^n$, $x^\intercal Ax$ is nonnegative,
    \item for some matrix $U$, $A = U^\intercal U$,
    \item $A$ is a nonnegative linear combination of matrices of the type $xx^\intercal$.
  \end{enumerate}
\end{theorem}

@@

\begin{theorem}[Proposition 5.1 of \cite{Lovasz95}]
  The minimum angle $\phi$ of any orthogonal representation of $G$ is given by $\cos^2\phi = 1/\vartheta(G)$.
\end{theorem}
\TODO{mark all theorems everywhere accordingly}

@@

  The running time of \cref{alg:maxStableSet} is polynomial, because we construct $|V|$ auxiliary graphs, each requiring calculating $\alpha$ once plus additional $O(|V|^2)$ time for constructing the graph.
\end{proof}

@@

There are at most $|V|$ maximum cardinality cliques in $G$. Adding a single clique to the list of maximum cardinality cliques requires constructing auxiliary graph for weighted maximum stable set, which is of size $O(|V|^2)$ and running \cref{alg:maxStableSet} on it. Therefore total running time is polynomial.

@@

Note that coloring $C'$ uses $\chi(G \setminus S) = \omega(G \setminus S) = \omega(G) -1$. We will call recursion at most $O(|V|)$ times, each step of recursion is polynomial in time. Therefore the running time of \cref{alg:minColoring} is polynomial.

>>>>

A natural problem for perfect graphs is a problem of coloring them. In 1988 Gr\"otschel et al. published an ellipsoid-method-based polynomial algorithm for coloring perfect graphs \cite{Grtschel1993}. We consider it in \Cref{sec:coloringEllipsoid}. However due to its use of the ellipsoid method this algorithm has been usually considered unpractical \cite{coloringSquareFree,Chudnovsky2003, coloringArtemis}.

There has been much progress on the quest of finding a more classical algorithm coloring perfect graphs, without the use of ellipsoid method (see \Cref{sec:classicalColoring}), however there is still no known polynomial combinatorial algorithm to do this. \todo{better wording of this paragraph}

@@

\Cref{sec:InformationTheory} and \Cref{sec:computingTheta} are based on lecture notes by Lovász \cite{Lovasz95}.

@@

Unfortunately, it is not known whether $\Theta(G)$ can be computed for all graphs in finite time. If we could calculate $\alpha(G^k)$ for a first few values of $k$ (we will show how to do it in \Cref{alg:maxStableSet}) we could have a lower bound on $\Theta(G)$. Let us now turn into search for some usable upper bound.

@@

\begin{theorem}[Lovász "sandwich theorem" \cite{Knuth1994}]
  \label{thm:sandwich}
  For any graph $G$:
  $$ \omega(G) \leq \vartheta(\overline{G}) \leq \chi(G).$$
\end{theorem}

@@

First, let us recall some definitions, with \cite{gilbertstrang2020} as a reference for the linear algebra.

@@

We have equivalent definitions of semidefinite matrices.
\begin{theorem}[Proposition 2.2 of \cite{Lovasz95}]
  For a real symmetric $n \times n$ matrix $A$, the following are equivalent:
  \begin{enumerate}[(i)]
    \item $A$ is positive semidefinite,
    \item \label{en:ei2} for every $x \in \mathbb{R}^n$, $x^\intercal Ax$ is nonnegative,
    \item for some matrix $U$, $A = U^\intercal U$,
    \item $A$ is a nonnegative linear combination of matrices of the type $xx^\intercal$.
  \end{enumerate}
\end{theorem}

@@

\begin{theorem}[Proposition 5.1 of \cite{Lovasz95}]
  The minimum angle $\phi$ of any orthogonal representation of $G$ is given by $\cos^2\phi = 1/\vartheta(G)$.
\end{theorem}

@@

  The running time of \Cref{alg:maxStableSet} is polynomial, because we construct $|V|$ auxiliary graphs, each requiring calculating $\alpha$ once plus additional $O(|V|^2)$ time for constructing the graph.
\end{proof}

@@

There are at most $|V|$ maximum cardinality cliques in $G$. Adding a single clique to the list of maximum cardinality cliques requires constructing auxiliary graph for weighted maximum stable set, which is of size $O(|V|^2)$ and running \Cref{alg:maxStableSet} on it. Therefore total running time is polynomial.

@@

Note that coloring $C'$ uses $\chi(G \setminus S) = \omega(G \setminus S) = \omega(G) -1$. We will call recursion at most $O(|V|)$ times, each step of recursion is polynomial in time. Therefore the running time of \Cref{alg:minColoring} is polynomial.

### 2020-09-15 paper/experiments.tex
Small fixes

An \emph{Erd\H{o}s--Rényi} model is a method of generating random graphs. Every possible edge is included in a graph with probability $1/2$. We generated them as long as we found enough perfect graphs for each $|V|$ we wanted to test. \todo{implement generating perfect graphs by removing long cycles} This method of generating perfect graphs is very inefficient and we couldn't generate any graphs with $|V| \geq 20$. Second, those graphs favour the na\"ive algortihm, because there is very low probability of long paths to appear.

Next, we generate bipartite graphs in similar manner -- each edge between two equal sized node sets has a probability of $1/2$ to appear. We also generate line graphs of bipartite graphs. This gives us graphs of varying $|V|$, so we repeated the process until we had sufficieng number of graphs for each $|V|$.

@@

\pagebreak
\section{Coloring Berge Graphs}

@@

Calculating $\vartheta(G)$ is the most complicated part of the coloring algorithm. With that done by an external library, the rest of the program is a straightforward implementation of the algorithms in \cref{sec:coloringEllipsoid}. In most of our tests, the majority of running time was consumed on calculating $\vartheta$ of various graphs. As it was done by an external library, there isn't much optimization potential for us. One thing of note is, that specifying the precision of $\vartheta$ we want to calculate to be $1/3$ sped up the algorithm by TODO. Our main goal of the implementation was to check if this method is still impractical, even on modern equipment. Let us proceed straight to experiments and results.

>>>>

% An \emph{Erd\H{o}s--Rényi} model is a method of generating random graphs. Every possible edge is included in a graph with probability $1/2$. We generated them as long as we found enough perfect graphs for each $|V|$ we wanted to test. \todo{implement generating perfect graphs by removing long cycles} This method of generating perfect graphs is very inefficient and we couldn't generate any graphs with $|V| \geq 20$. Second, those graphs favour the na\"ive algortihm, because there is very low probability of long paths to appear.

We generate random bipartite graphs by creating two equal-sized sets of vertices and every possible edge between them has a probability of $1/2$ to appear. We also generate line graphs of bipartite graphs. This gives us graphs of varying $|V|$, so we repeated the process until we had sufficieng number of graphs for each $|V|$.

@@

\section{Coloring Berge Graphs}

@@

Calculating $\vartheta(G)$ is the most complicated part of the coloring algorithm. With that done by an external library, the rest of the program is a straightforward implementation of the algorithms in \Cref{sec:coloringEllipsoid}. In most of our tests, the majority of running time was consumed on calculating $\vartheta$ of various graphs. As it was done by an external library, there isn't much optimization potential for us. One thing of note is, that specifying the precision of $\vartheta$ we want to calculate to be $1/3$ sped up the algorithm by TODO. Our main goal of the implementation was to check if this method is still impractical, even on modern equipment. Let us proceed straight to experiments and results.

### 2020-09-15 paper/implementation.tex
Small fixes

\section{Berge graphs recognition: na\"ive approach}
When implementing an algorithm with the running time of $O(n^9)$ (especially one as complex as CCLSV), one of the first questions that pop up is whether this algorithm is applicable at all. We couldn't find any existing programs that test whether the graph is perfect, be it an implementation of CCLSV, or a na\"ive approach.\todo{well.. there's this java impl} So, to test the usability of CCLSV implementation, we measured it against our own na\"ive Perfect recognition algorithm. Let us describe it briefly.

We will try find an odd hole directly, then, we run the same algorithm on $\overline{G}$, and if no odd hole is found we report that the graph is Perfect, else that it is not.

To find an odd hole, we use a backtracking technique. We will enumerate all paths and holes and for each test if it is an odd hole. A partial solution to our problem is a path in $G$. To move forward from a partial solution, we append consecutively to its end all neighbors of current path's last vertex and continue recursively after each. When a current solution is an odd hole we return it. When we exhaust all possible paths from the current partial solution, we remove last vertex and continue.

This backtracking algorithm is very simple, but has a great potential for optimizations in implementation. We also use path enumeration algorithm in CCLSV algorithm, so we give much attention to its optimization.

@@

We notice that returning a new path for each call of the method is wasteful. Therefore, we need a method that receives a reference to a path on the input (or a special flag indicating it should generate first path) and returns next path in some order (or a first path, or a code signaling all paths have been generated) using memory from the input. As this method will be used many, many times we require of it to work in place with constant additional space.

@@

\paragraph{Other uses for path generation}
\label{sec:usesGeneration}

@@

With all those optimizations made, our na\"ive algorithm proved to be quite fast in some cases, although its running time is very dependent on the properties of the input, as we should expect from an optimized non-polynomial algorithm. See \cref{sec:experiments} for running times.

@@

The Berge recognition algorithm's running time is $O(n^9)$, which brings into question its applicability to any real use case. Although time complexity is indeed a limiting factor, a number of lower level optimizations done on implementation's level (\cref{sec:Optimizations}) make a very big difference and make it more viable than na\"ive algorithm, at least on some test cases. Also, we explore a new frontier of implementing its most time consuming part on massively parallel GPU architecture, with some good results (\cref{sec:CUDA}).

@@

Whole algorithm was divided into subroutines used in many places. One of such subroutines is a path generation algorithm described above (\cref{sec:usesGeneration}). There are many other generalized methods we implemented: checking if a set of vertices is $X$-complete, getting a set of components of a graph, creating a induced graph or finding a shortest path, such that each internal vertex satisfies a predicate. Each of those and many more methods have unit tests that check their general use and edge cases. This allowed us to debug very effectively and have complex algorithms be simple to analyze.

@@

In addition to unit tests, we employ a range of end to end tests, that check the final answer of the algorithm. We compare answer to the na\"ive algorithm's answer and also to the result of the algorithm on CUDA (\cref{sec:CUDA}).

@@

To identify parts of CCLSV to implement, we used callgrind and found potential bottlenecks. The initial tests showed that with growing size of a graph some parts take more and more relative time. Because of callgrind's slow execution times compared to just running the program, we used manual timers for bigger tests. Using this method we identified testing all possible near-cleaners (\cref{alg:testNearCleaner}) as the biggest bottleneck for larger graphs. We considered two approaches: run whole \cref{alg:testNearCleaner} on a single CUDA core, testing multiple $X$s in parallel, or parallelize \cref{alg:testNearCleaner} itself and run it on one $X$ at a time. It turned out that second approach is better and much simpler.

Let us recall the \cref{alg:testNearCleaner}. It calculates array $R$ of shortest paths, then for each 3-vertex path and an additional vertex it does $O(1)$ checks to see if they along with two paths from $R$ give us an odd hole. We will parallelise the work after calculating $R$. Let us notice, that for all $X$s all 3-vertex paths are the same. We calculate them beforehand, then each thread receives a 3-vertex path $x_1 - x_2 - x_3$ and an additional vertex $y_1$ and performs the required checks. This is almost perfect scenario for the GPU -- we do a simple SIMD work in parallel, without a lot scattered memory access.

>>>>

\section{Berge graphs recognition: na\"ive approach}
When implementing an algorithm with the running time of $O(n^9)$ (especially one as complex as CCLSV), one of the first questions that pop up is whether this algorithm is applicable at all. We couldn't find any existing programs that test whether the graph is perfect, be it an implementation of CCLSV, or a na\"ive approach.\todo{well.. there's this java impl} So, to test the usability of CCLSV implementation, we measured it against our own na\"ive perfect recognition algorithm. Let us describe it briefly.

We try find an odd hole directly, then, we run the same algorithm on $\overline{G}$, and if no odd hole is found we report that the graph is Perfect, else that it is not.

To find an odd hole, we use a backtracking technique. We will enumerate all paths and holes and for each test if it is an odd hole. A partial solution to our backtracking problem is a path in $G$. To move forward from a partial solution, we append consecutively to its end all neighbors of current path's last vertex and continue recursively after each. When a current solution is an odd hole we return it. When we exhaust all possible paths from the current partial solution, we remove last vertex and continue.

This backtracking algorithm is very simple, but has a great potential for optimizations in implementation. We also use path enumeration algorithm in the CCLSV algorithm, so we give much attention to its optimization.

@@

We will create a method that enumerates all paths in some order. We notice that returning a new path for each call of the enumeration method is wasteful. Therefore, we need a method that receives a reference to a path on the input (or a special flag indicating it should generate first path) and returns next path in some order (or a first path, or a code signaling all paths have been generated) using memory from the input. As this method will be used many, many times we require of it to work in place with constant additional space.

@@

\paragraph{Other uses for path generation.}
\label{sec:usesGeneration}

@@

With all those optimizations made, our na\"ive algorithm proved to be quite fast in some cases, although its running time is very dependent on the properties of the input, as we should expect from an optimized non-polynomial algorithm. See \Cref{sec:experiments} for running times.

@@

The Berge recognition algorithm's running time is $O(n^9)$, which brings into question its applicability to any real use case. Although time complexity is indeed a limiting factor, a number of lower level optimizations done on implementation's level (\Cref{sec:Optimizations}) make a very big difference and make it more viable than na\"ive algorithm, at least on some test cases. Also, we explore a new frontier of implementing its most time consuming part on massively parallel GPU architecture, with some good results (\Cref{sec:CUDA}).

@@

Whole algorithm was divided into subroutines used in many places. One of such subroutines is a path generation algorithm described above (\Cref{sec:usesGeneration}). There are many other generalized methods we implemented: checking if a set of vertices is $X$-complete, getting a set of components of a graph, creating a induced graph or finding a shortest path, such that each internal vertex satisfies a predicate. Each of those and many more methods have unit tests that check their general use and edge cases. This allowed us to debug very effectively and have complex algorithms be simple to analyze.

@@

In addition to unit tests, we employ a range of end to end tests, that check the final answer of the algorithm. We compare answer to the na\"ive algorithm's answer and also to the result of the algorithm on CUDA (\Cref{sec:CUDA}).

@@

To identify parts of CCLSV to implement, we used callgrind and found potential bottlenecks. The initial tests showed that with growing size of a graph some parts take more and more relative time. Because of callgrind's slow execution times compared to just running the program, we used manual timers for bigger tests. Using this method we identified testing all possible near-cleaners (\Cref{alg:testNearCleaner}) as the biggest bottleneck for larger graphs. We considered two approaches: run whole \Cref{alg:testNearCleaner} on a single CUDA core, testing multiple $X$s in parallel, or parallelize \Cref{alg:testNearCleaner} itself and run it on one $X$ at a time. It turned out that second approach is better and much simpler.

Let us recall the \Cref{alg:testNearCleaner}. It calculates array $R$ of shortest paths, then for each 3-vertex path and an additional vertex it does $O(1)$ checks to see if they along with two paths from $R$ give us an odd hole. We will parallelise the work after calculating $R$. Let us notice, that for all $X$s all 3-vertex paths are the same. We calculate them beforehand, then each thread receives a 3-vertex path $x_1 - x_2 - x_3$ and an additional vertex $y_1$ and performs the required checks. This is almost perfect scenario for the GPU -- we do a simple SIMD work in parallel, without a lot scattered memory access.

### 2020-09-15 paper/perfectGraphs.tex
Small fixes

The most fundamental problem -- the problem of recognizing perfect graphs -- was open since its posing in 1961 until recently. Its solution, a polynomial algorithm recognizing perfect graphs is a union of the strong perfect graph theorem (\cref{sec:SPGT}) stating that a graph is perfect if and only if it is Berge and an algorithm for recognizing Berge graphs in polynomial time (\cref{sec:recognizingBerge}).

@@

Perfect graphs are interesting not only because of their theoretical properties, but they are also used in other areas of study e.g. integrality of polyhedra \cite{Chvtal1975, Chudnovsky2003}, radio channel assignment problem \cite{McDiarmid99, McDiarmid2000} and appear in the theory of Shannon capacity of a graph \cite{Lovasz1979}. Also, as pointed out in \cite{alfonsinPerfect2001, Chudnovsky2003} algorithms to solve semi-definite programs grew out of the theory of perfect graphs. We will take a look at semi-definite programs and at perfect graph's relation to Shannon capacity in \cref{sec:ShannonCapacity}.

@@

Then, since $\alpha(H) = \omega(\overline{H})$ and $\omega(H) = \alpha(\overline{H})$ \cref{thm:omegaalpha} implies \cref{thm:pgt}.

Odd holes are not perfect, since their chromatic number is 3 and their largest cliques are of size 2. It is also easy to see, that an odd antihole of size $n$ has a chromatic number of $\frac{n+1}{2}$ and largest cliques of size $\frac{n-1}{2}$. A graph with no odd hole and no odd antihole is called \emph{Berge} (\cref{def:bergeGraph}) after Claude Berge who studied perfect graphs.

@@

\subsubsection{Basic classes of perfect graphs}
Bipartite graphs are perfect, since we can color them with two colors. From the theorem of König we get that line graphs of bipartite graphs are also perfect \cite{Knig1916, GC03}. From the perfect graph theorem (\cref{thm:pgt}) it follows that complements of bipartite graphs and complement of line graphs of bipartite graphs are also perfect. We will call these four classes \emph{basic}.

@@

The proof of \cref{thm:spgt} is a consequence of the \emph{Decomposition theorem}:
\begin{theorem}[Decomposition theorem, stated as Theorem 4.2 of \cite{GC03}]
	\label{thm:decomposition}
	Every Berge graph $G$ is basic or has a skew partition or a homogeneous pair, or either $G$ or $\overline{G}$ has a 2-join.
\end{theorem}

See \cite{MC06} for the full proof of \cref{thm:spgt} and \cite{GC03} for its shorter overview.

@@

The following is based on the paper by Maria \citeauthor{MC05} \citetitle*[]{MC05} \cite{MC05}. We will not provide full proof of its correctness, but will aim to show the intuition behind the algorithm.

Berge graph recognition algorithm (later called CCLSV from the names of its authors) could be divided into two parts: first we check if either $G$ or $\overline{G}$ contain any of a number of simple forbidden structures (\cref{SimpleStructures}). If they do, we output that graph is not Berge and stop. Else, we check if there is a near-cleaner for a shortest odd hole (\cref{AmenableHoles}).

@@

\begin{wrapfigure}{r}{0.35\textwidth}
	\input{tikzpictures/pyramid.tex}
	\caption{An example  of a pyramid.}%
	\vspace{-1.2cm}
\end{wrapfigure}

A \emph{pyramid} in G is an induced subgraph formed by the union of a triangle \footnote{A triangle is a clique $K_3$.} $\{b_1,b_2,b_3\}$, three paths $\{P_1, P_2, P_3\}$ and another vertex $a$, so that:
\begin{itemize}
	\item $\forall_{1\leq i \leq 3}$ $P_i$ is a path between $a$ and $b_i$,
	\item $\forall_{1\leq i < j \leq 3}$ $a$ is the only vertex in both $P_i$ and $P_j$ and $b_ib_j$ is the only edge between $V(P_i)\setminus\{a\}$ and $V(P_j)\setminus\{a\}$,
	\item $a$ is adjacent to at most one of $\{b_1, b_2, b_3\}$.
\end{itemize}

@@

\paragraph{Finding Pyramids}

@@

% \begin{wrapfigure}[7]{r}{0.35\textwidth}
\begin{wrapfigure}{r}{0.35\textwidth}
	\input{tikzpictures/jewel.tex}
	\caption{An example of a jewel.}
	\vspace{-1.5cm}
\end{wrapfigure}

@@

A configuration of type $\T_1$ is a hole of length 5. To find it, we simply iterate over all paths of length 4 and check if there exists a fifth vertex to complete the hole. See \cref{sec:usesGeneration} for more implementation details.

@@

\begin{theorem}[Theorem 6.3 of \cite{MC05}]
	If $G$ contains configuration of type $\T_2$ then $G$ is not Berge.
\end{theorem}
\begin{proof}
	Let $(v_1, v_2, v_3, v_4, P, X)$ be a configuration of type $\T_2$. Let us assume that $G$ is not Berge and consider the following:
	\begin{itemize}
		\item if $P$ is even, then $v_1, v_2, v_3, v_4, P, v_1$ is an odd hole,
		\item if $P$ is of length 3. \todo{I merged a couple of proofs from \cite{MC06}, check in the morning if this is correct.} Let us name its vertices $v_1, p_2, p_3, v_4$. It follows from \cref{lem:Roussel-Rubio}, that there exists an odd antipath between $p_2$ and $p_3$ with interior in $X$. We can complete it with $v_2p_2$ and $v_2p_3$ into an odd antihole,
		\item if $P$ is odd with the length of at least 5 \todo{check in the morning}, it follows from \cref{lem:Roussel-Rubio} that we have $x, y \in X$ with only two edges to $P$ being $xp_2$ and $yp_{n-1}$. This gives us an odd hole: $v_2, x, p_2, \ldots, p_{n-1}, y, v_2$.
	\end{itemize}
\end{proof}

@@

	\noindent Output: Returns whether $G$ contains a configuration of type $T_3$ as an induced subgraph.
\end{alg}

@@

To prove the correctness of \cref{alg:listNearCleaners} we will need the following theorem.

@@

\begin{theorem}[Theorem 9.2 of \cite{MC05}]
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Then the result of the \Cref{alg:listNearCleaners} on the graph $G$ is a set $\mathcal{R}$, such that for every shortest amenable hole $C$, there is a near-cleaner for $C$ in $\mathcal{R}$.
\end{theorem}
\begin{proof}
	Let us suppose that $C$ is an amenable hole in $G$. By \cref{thm:91}, there is a relevant triple $(a, b, c)$ satisfying that theorem. Let $T$ be the set of all $C$-major vertices not in $X(a,b,c)$. From \cref{thm:91} we get that $T$ is anticonnected. Since $C$ is amenable, there is an edge $uv$ of $C$ that is $T$-complete, and therefore $T \subseteq N(u, v)$. But then $N(u, v) \cup X(a, b, c) \in \mathcal{R}$ is a near-cleaner for $C$. Therefore the output of the \cref{alg:listNearCleaners} is correct.
\end{proof}

@@

Below we will prove that if the \cref{alg:testNearCleaner} reports an odd hole in $G$, there indeed is one. The proof of the correctness of the other possible result is more complicated, see section 4 and theorem 5.1 of \cite{MC05}.

\begin{theorem}[Theorem 5.1 of \cite{MC05}]
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Let $X$ be a subset of $V(G)$. If \Cref{alg:testNearCleaner} reports an odd hole, then $G$ contains an odd hole.
\end{theorem}
\begin{proof}
	Let us suppose that there is a choice of $x_1$, $x_2$, $x_3$, $y_1$ satisfying the three conditions in the \cref{alg:testNearCleaner} and let $y_2$ and $n$ be defined as in there. We claim that $G$ contains an odd hole.

	{\makeatletter
	\let\par\@@par
	\par\parshape0
	\everypar{}\begin{wrapfigure}{r}{0.3\textwidth}
		\input{tikzpictures/nearCleanerFound.tex}
		\caption{An odd hole is found}%
		% \vspace{-0.2cm}
		\end{wrapfigure}
		Let $R(x_1, y_1) = p_1-\ldots -p_n$, and let $R(x_2, y_1) = q_1-\ldots -q_{n+1}$, where $p_1 = x_1$, $p_n = q_{n+1} = y_1$, $q_1 = x_2$ and $q_n = y_2$. From the definition of $R(x_1, y_1)$ and $R(x_2, y_1)$, none of $p_2, \ldots, p_{n-1}, q_2, \ldots, q_n$ belong to $X$. Also, from the definition of $y_1$, $y_1 \notin X$.

@@

\begin{wrapfigure}{r}{0.35\textwidth}

\end{wrapfigure}

\subsection{Summary}

\TODO{todo?}

>>>>

The most fundamental problem -- the problem of recognizing perfect graphs -- was open since its posing in 1961 until recently. Its solution, a polynomial algorithm recognizing perfect graphs is a union of the strong perfect graph theorem (\Cref{sec:SPGT}) stating that a graph is perfect if and only if it is Berge and an algorithm for recognizing Berge graphs in polynomial time (\Cref{sec:recognizingBerge}).

@@

Perfect graphs are interesting not only because of their theoretical properties, but they are also used in other areas of study e.g. integrality of polyhedra \cite{Chvtal1975, Chudnovsky2003}, radio channel assignment problem \cite{McDiarmid99, McDiarmid2000} and appear in the theory of Shannon capacity of a graph \cite{Lovasz1979}. Also, as pointed out in \cite{alfonsinPerfect2001, Chudnovsky2003} algorithms to solve semi-definite programs grew out of the theory of perfect graphs. We will take a look at semi-definite programs and at perfect graph's relation to Shannon capacity in \Cref{sec:ShannonCapacity}.

@@

Then, since $\alpha(H) = \omega(\overline{H})$ and $\omega(H) = \alpha(\overline{H})$ \Cref{thm:omegaalpha} implies \Cref{thm:pgt}.

Odd holes are not perfect, since their chromatic number is 3 and their largest cliques are of size 2. It is also easy to see, that an odd antihole of size $n$ has a chromatic number of $\frac{n+1}{2}$ and largest cliques of size $\frac{n-1}{2}$. A graph with no odd hole and no odd antihole is called \emph{Berge} (\Cref{def:bergeGraph}) after Claude Berge who studied perfect graphs.

@@

\subsubsection{Basic classes of perfect graphs}
Bipartite graphs are perfect, since we can color them with two colors. From the theorem of König we get that line graphs of bipartite graphs are also perfect \cite{Knig1916, GC03}. From the perfect graph theorem (\Cref{thm:pgt}) it follows that complements of bipartite graphs and complement of line graphs of bipartite graphs are also perfect. We will call these four classes \emph{basic}.

@@

The proof of \Cref{thm:spgt} is a consequence of the \emph{Decomposition theorem}:
\begin{theorem}[Decomposition theorem, stated as Theorem 4.2 of \cite{GC03}]
	\label{thm:decomposition}
	Every Berge graph $G$ is basic or has a skew partition or a homogeneous pair, or either $G$ or $\overline{G}$ has a 2-join.
\end{theorem}

See \cite{MC06} for the full proof of \Cref{thm:spgt} and \cite{GC03} for its shorter overview.

@@

The following is based on the paper by \citeauthor{MC05} \citetitle*[]{MC05} \cite{MC05}. We will not provide full proof of its correctness, but will aim to show the intuition behind the algorithm.

Berge graph recognition algorithm (later called CCLSV from the names of its authors) could be divided into two parts: first we check if either $G$ or $\overline{G}$ contain any of a number of simple forbidden structures (\Cref{SimpleStructures}). If they do, we output that graph is not Berge and stop. Else, we check if there is a near-cleaner for a shortest odd hole (\Cref{AmenableHoles}).

@@

A \emph{pyramid} in G is an induced subgraph formed by the union of a triangle \footnote{A triangle is a clique $K_3$.} $\{b_1,b_2,b_3\}$, three paths $\{P_1, P_2, P_3\}$ and another vertex $a$, so that:

\begin{minipage}{0.60\textwidth}
\begin{itemize}
	\item $\forall_{1\leq i \leq 3}$ $P_i$ is a path between $a$ and $b_i$,
	\item $\forall_{1\leq i < j \leq 3}$ $a$ is the only vertex in both $P_i$ and $P_j$ and $b_ib_j$ is the only edge between $V(P_i)\setminus\{a\}$ and $V(P_j)\setminus\{a\}$,
	\item $a$ is adjacent to at most one of $\{b_1, b_2, b_3\}$.
\end{itemize}
\end{minipage}%
% \begin{minipage}
\begin{wrapfigure}{r}{0.32\textwidth}
	\vspace{-2.5cm}
	\input{tikzpictures/pyramid.tex}
	\caption{An example  of a pyramid.}%
	\vspace{-1.2cm}
\end{wrapfigure}
% \end{minipage}

@@

\paragraph{Finding Pyramids.}

@@

% \begin{wrapfigure}[7]{r}{0.35\textwidth}
\begin{wrapfigure}{r}{0.35\textwidth}
	\input{tikzpictures/jewel.tex}
	\caption{An example of a jewel.}
	\vspace{-1.8cm}
\end{wrapfigure}

@@

A configuration of type $\T_1$ is a hole of length 5. To find it, we simply iterate over all paths of length 4 and check if there exists a fifth vertex to complete the hole. See \Cref{sec:usesGeneration} for more implementation details.

@@

\begin{theorem}[Theorem 6.3 of \cite{MC05}]
	If $G$ contains configuration of type $\T_2$ then $G$ is not Berge.
\end{theorem}
\begin{proof}
	Let $(v_1, v_2, v_3, v_4, P, X)$ be a configuration of type $\T_2$. Let us assume that $G$ is not Berge and consider the following:
	\begin{itemize}
		\item if $P$ is even, then $v_1, v_2, v_3, v_4, P, v_1$ is an odd hole,
		\item if $P$ is of length 3. \todo{I merged a couple of proofs from \cite{MC06}, check in the morning if this is correct.} Let us name its vertices $v_1, p_2, p_3, v_4$. It follows from \Cref{lem:Roussel-Rubio}, that there exists an odd antipath between $p_2$ and $p_3$ with interior in $X$. We can complete it with $v_2p_2$ and $v_2p_3$ into an odd antihole,
		\item if $P$ is odd with the length of at least 5 \todo{check in the morning}, it follows from \Cref{lem:Roussel-Rubio} that we have $x, y \in X$ with only two edges to $P$ being $xp_2$ and $yp_{n-1}$. This gives us an odd hole: $v_2, x, p_2, \ldots, p_{n-1}, y, v_2$.
	\end{itemize}
\end{proof}

@@

	\noindent Output: Returns whether $G$ contains a configuration of type $\T_3$ as an induced subgraph.
\end{alg}

@@

To prove the correctness of \Cref{alg:listNearCleaners} we will need the following theorem.

@@

\begin{theorem}[Theorem 9.2 of \cite{MC05}]
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Then the result of the \Cref{alg:listNearCleaners} on the graph $G$ is a set $\mathcal{R}$, such that for every shortest amenable hole $C$, there is a near-cleaner for $C$ in $\mathcal{R}$.
\end{theorem}
\begin{proof}
	Let us suppose that $C$ is an amenable hole in $G$. By \Cref{thm:91}, there is a relevant triple $(a, b, c)$ satisfying that theorem. Let $T$ be the set of all $C$-major vertices not in $X(a,b,c)$. From \Cref{thm:91} we get that $T$ is anticonnected. Since $C$ is amenable, there is an edge $uv$ of $C$ that is $T$-complete, and therefore $T \subseteq N(u, v)$. But then $N(u, v) \cup X(a, b, c) \in \mathcal{R}$ is a near-cleaner for $C$. Therefore the output of the \Cref{alg:listNearCleaners} is correct.
\end{proof}

@@

Below we will prove that if the \Cref{alg:testNearCleaner} reports an odd hole in $G$, there indeed is one. The proof of the correctness of the other possible result is more complicated, see section 4 and theorem 5.1 of \cite{MC05}.

\begin{theorem}[Theorem 5.1 of \cite{MC05}]
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Let $X$ be a subset of $V(G)$. If \Cref{alg:testNearCleaner} reports an odd hole, then $G$ contains an odd hole.
\end{theorem}
\begin{proof}
	Let us suppose that there is a choice of $x_1$, $x_2$, $x_3$, $y_1$ satisfying the three conditions in the \Cref{alg:testNearCleaner} and let $y_2$ and $n$ be defined as in there. We claim that $G$ contains an odd hole.

	{\makeatletter
	\let\par\@@par
	\par\parshape0
	\everypar{}\begin{wrapfigure}{r}{0.35\textwidth}
		\input{tikzpictures/nearCleanerFound.tex}
		\caption{An odd hole is found}%
		\vspace{-0.3cm}
		\end{wrapfigure}
		Let $R(x_1, y_1) = p_1-\ldots -p_n$, and let $R(x_2, y_1) = q_1-\ldots -q_{n+1}$, where $p_1 = x_1$, $p_n = q_{n+1} = y_1$, $q_1 = x_2$ and $q_n = y_2$. From the definition of $R(x_1, y_1)$ and $R(x_2, y_1)$, none of $p_2, \ldots, p_{n-1}, q_2, \ldots, q_n$ belong to $X$. Also, from the definition of $y_1$, $y_1 \notin X$.

@@

% \subsection{Summary}

% \TODO{todo?}
